---
title: "Histograma"
output: html_notebook
---

Histrograma: es una gráfica que representa la acumulación de valores de una variable, indicando en qué zonas existe una mayor densidad de ocurrencias. Se disponen en rectángulos y pueden ser representados junto a una curva de densidad que completa la gráfica.

Load some data to work with:
```{r}
data("ChickWeight")
?hist
head(ChickWeight)
```


Simple example
```{r}
hist(ChickWeight$weight)
```



changing the pattern of the histogram bars
```{r}
hist(ChickWeight$weight, density = 30)
```



Control where the histogram breaks occur
```{r}
hist(ChickWeight$weight, density = 30, breaks = c(0,100,200,max(ChickWeight$weight)))
```



Another example of controlling the breaks
```{r}
hist(ChickWeight$weight, breaks = fivenum(ChickWeight$weight))
```





More on Histograms:
pacman must already be installed; then load contributed
packages (including pacman) with pacman
```{r}
pacman::p_load(pacman, rio, tidyverse)
```

Load data to work with:
Using the diamonds dataset from ggplot2
```{r}
?diamonds  # Get information about the diamonds dataset
diamonds   # Display the beginning of the dataset
```



Histogram with defaults
```{r}
hist(diamonds$price)
```



Histogram with options
```{r}
hist(diamonds$price,
  breaks = 7,  # Suggest number of breaks
  main   = "Histogram of Price of Diamonds",
  sub    = "(Source: ggplot2::diamonds)",
  ylab   = "Frequency",
  xlab   = "Price of Diamonds",
  border = TRUE,  # No borders on bars
  col    = "#CD0000"  # red3
  )
```



CLEAN UP 

Clear environment
```{r}
rm(list = ls()) 
```

Clear packages
```{r}
p_unload(all)  # Remove all add-ons
```

Clear plots
```{r}
dev.off()  # But only if there IS a plot
```

Clear console
```{r}
cat("\014")  # ctrl+L
```

Clear mind :)
