---
title: "R Gráficos de caja o boxplot"
output: html_notebook
---

Gráficos de caja o boxplot: Los gráficos de caja o bigote permiten visualizar dónde se encuentra el 50 % de los datos de una variable indicando el rango de valores central que queda delimitado por el rectángulo o caja. Además, generan un rango razonable donde cabe esperar que se encuentren las lecturas de la misma y marcan como v alo r e s a típic o s los que se salen de ese rango.

 - The bottom line represents the minimum value. 
 - The box encompasses the second quartile to the third quartile. 
 - The dark black line is the median and 
 - The top line is the fourth quartile. 
And then the circles above the top line are outliers. 

```{r}
?boxplot # Produce box-and-whisker plot(s) of the given (grouped) values.
```


Quick intro using 
```{r}
data("mtcars")
head(mtcars)
```

Plotting a simple example with just 1 atribute
```{r}
boxplot(mtcars$mpg)
```

The numbers used to create the boxplot
```{r}
fivenum(mtcars$mpg) 
```

Comparing one value to another with alternative notation, mpg as a function of cyl:
```{r}
boxplot(mpg ~ cyl, data = mtcars, col = "lightgray", varwidth = TRUE, 
        main = "mpg vs cylinders",
        ylab = "mpg",xlab = "cylinders")
```


More on boxplots
pacman must already be installed; then load contributed
packages (including pacman) with pacman
```{r}
pacman::p_load(pacman, tidyverse)
```

LOAD DATA 
Using the diamonds dataset from ggplot2
```{r}
?diamonds  # Get information about the diamonds dataset
diamonds   # Display the beginning of the dataset
```



Boxplot with defaults
```{r}
boxplot(diamonds$price)
```



Similar process using pipes
```{r}
diamonds %>%
  select(price) %>%
  boxplot()
```



Boxplot with options
```{r}
diamonds %>%
  select(price) %>%
  boxplot(
    horizontal = T,  # Horizontal
    notch  = T,      # Confidence interval for median
    main   = "Boxplot of Price of Diamonds",
    sub    = "(Source: ggplot2::diamonds)",
    xlab   = "Price of Diamonds",
    col    = "#CD0000"  # red3
  )
```



BOXPLOTS BY GROUP 
Boxplots by group using plot()
```{r}
diamonds %>%
  select(color, price) %>%
  plot()
```



Boxplots by group using boxplot()
```{r}
diamonds %>%
  select(color, price) %>%
  boxplot(
    price ~ color,  # Tilde indicates formula
    data  = . ,     # Dot is placeholder for pipe
    col   = "#CD0000"
  )
```





CLEAN UP 

Clear environment
```{r}
rm(list = ls()) 
```

Clear packages
```{r}
p_unload(all)  # Remove all add-ons
```

Clear plots
```{r}
dev.off()  # But only if there IS a plot
```

Clear console
```{r}
cat("\014")  # ctrl+L
```

Clear mind :)
