---
title: "R Time Series"
output: html_notebook
---

Gráfico de serie temporal: Una serie temporal es una cadena de valores tomados en distintos instantes. Representarlas es útil para analizar tendencias y puntos de cambio.

Studies of time-series data involve 2 fundamental questions:
 - waht happened? description
 - what will happen next? forecasting
To work with a time series in R, it is needed to place it into a time-series object. 
A vector of numbers, or a column in a data frame, can saved as a time-series using the ts() function:

timeseries <- ts(data, start=, end=, frequency)

 - data: is a numeric vector/column in a dataframe containing the observations
 - start: specifies the series start time
 - end: specifies the end time (optional)
 - frequency: indicates the number of observations per unit time
 
Example:
```{r}
sales <- c(18,33,41,7,34,35,24,25,24,21,25,20,22,31,40,29,25,21,22,54,31,25,26,35)
```

```{r}
tsales <- ts(sales, start=c(2003, 1),frequency = 12)
tsales
```

```{r}
plot(tsales)
```

```{r}
start(tsales)
```

```{r}
end(tsales)
```

```{r}
frequency(tsales)
```

```{r}
tsales.subset <- window(tsales, start = c(2003,5), end=c(2004,6))
tsales.subset
```


More on plotting time-series
pacman must already be installed; then load contributed
packages (including pacman) with pacman
```{r}
pacman::p_load(pacman, tidyverse)
```


Load base packages manually
```{r}
library(datasets)  # For example datasets
```


SINGLE TIME SERIES 
uspop
```{r}
?uspop  # Get info about data
uspop   # Display data

?ts  # Get info about time-series objects
```


Plot with default plot()
```{r}
plot(uspop)
```


Plot with options
```{r}
uspop %>% 
  plot(
    main = "US Population 1790â1970 ",
    sub  = "(Source: datasets::uspop)",
    xlab = "Year",
    ylab = "Population (in millions)",
  )
abline(v = 1930, col = "lightgray")
text(1930, 10, "1930", col = "red3") 
abline(v = 1940, col = "lightgray")
text(1940, 2, "1940", col = "red3") 
```


Plot with ts.plot()
```{r}
?ts.plot
```

Although this can be used for a single time series, plo is easier to use and is preferred.
```{r}
ts.plot(uspop)
```


Plot with plot.ts()
More powerful alternative
```{r}
?plot.ts
plot.ts(uspop)
```



MULTIPLE TIME SERIES 

EuStockMarkets
```{r}
?EuStockMarkets
EuStockMarkets
```


Three different plot functions
```{r}
plot(EuStockMarkets)     # Stacked windows
plot.ts(EuStockMarkets)  # Identical
ts.plot(EuStockMarkets)  # One window
```




CLEAN UP 

Clear environment
```{r}
rm(list = ls()) 
```

Clear packages
```{r}
p_unload(all)  # Remove all add-ons
```

Clear plots
```{r}
dev.off()  # But only if there IS a plot
```

Clear console
```{r}
cat("\014")  # ctrl+L
```

Clear mind :)
