---
title: "rep() for repetitive vectors"
output: html_notebook
---

rep replicates the values in x. It is a generic function, and the (internal) default method is described here.

rep.int and rep_len are faster simplified versions for two common cases. Internally, they are generic, so methods can be defined for them.

rep(x, ...)

rep.int(x, times)

rep_len(x, length.out)

```{r}
?rep
rep("Hello", 5)
```

```{r}
rep("hello", length.out = 10)
rep_len("hello", length.out = 10)
```

```{r}
rep(c("hello","bob"), 3) #repeats the vector 3 times
rep(c("hello","bob"), each = 3) #repeats each vector element 3 times
```


rep() applied to dataframes returns a list
```{r}
df <- data.frame(x = c(4,5,6), y = c(7,8,9))
dfrep <- rep(df,3) 
```

How to set quarter identifiers for a random number of years:
```{r}
paste0(2003:2021,"-") # range of years
rep(paste0(2003:2021,"-"), 4) # 4x years - one for each quarter
paste0(rep(paste0(2003:2021,"-"),4),c("Q1","Q2","Q3","Q4"))
sort(paste0(rep(paste0(2003:2021,"-"),4),c("Q1","Q2","Q3","Q4"))) # sort by year
```

End
```{r}
rm(list=ls())
```
ctrl+l
