---
title: "plotting"
output: html_notebook
---

Quick intro using 
```{r}
data("ChickWeight")
```

Plotting 2 variables
```{r}
?plot
plot(ChickWeight$Time,ChickWeight$weight)
```

Generic X-Y Plotting

```{r}
plot(ChickWeight$Time,ChickWeight$weight,
     type = "s", # s for stair steps
     main = "plotting weight")
```

Adding a line on the plot
```{r}
plot(ChickWeight$Time,ChickWeight$weight,
     type = "l", # l for lines
     main = "plotting weight")
lines(c(1,20),c(350,50),
      col="red") # lines(c(x1-value,x2-value),c(y1-value,y2-value))
```


More on scatterplots:
pacman must already be installed; then load contributed
packages (including pacman) with pacman
```{r}
pacman::p_load(pacman, rio, tidyverse)
```

Load data to work with:
```{r}
df <- import("~/GitHub/Learning R/StateData.xlsx") %>%
  as_tibble() %>%
  select(state_code, 
    psychRegions,
    instagram:modernDance) %>% 
  mutate(psychRegions = as.factor(psychRegions)) %>%
  print()
```

Scatterplots matrix: showing all possible associations
```{r}
df %>%
  select(instagram:modernDance) %>% 
  plot()  # Plot all associations
```


Bivariate scatterplot with defaults / Diagrama de dispersión bivariante con valores predeterminados
Se seleccionan las variables scrapbook y modernDance (abajo a la derecha), muestra una covarianza negativa con outliers
```{r}
df %>%
  select(scrapbook:modernDance) %>% 
  plot()
```


Bivariate scatterplot with options / Diagrama de dispersiones bivariante con opciones:
```{r}
df %>%
  select(scrapbook:modernDance) %>% 
  plot(
    main = "Scatterplot of Google Searches by State",
    xlab = "Searches for \"scrapbook\"",
    ylab = "Searches for \"modern dance\"",
    col  = "gray",  # Color of points
    pch  = 20,      # "Plotting character" (small circle)
  )
df %>%
  lm(modernDance ~ scrapbook, data = .) %>%
  abline()
```


```{r}
?pch  # See plotting characters
?lm # lm is used to fit linear models. It can be used to carry out regression, single stratum analysis of variance and analysis of covariance (although aov may provide a more convenient interface for these).
?abline # This function adds one or more straight lines through the current plot
```

Add fit linear regression line (y ~ x) 
lm - linear model, generic command for regression
~ indicates one variable (modernDance - y axis) is a function of other (scrapbook - X axis)
%>% pipes from tidyverse
data = . means take the previous thing, which is df, and feed it there
```{r}
df %>%
  select(scrapbook:modernDance) %>% 
  plot(
    main = "Scatterplot of Google Searches by State",
    xlab = "Searches for \"scrapbook\"",
    ylab = "Searches for \"modern dance\"",
    col  = "gray",  # Color of points
    pch  = 20,      # "Plotting character" (small circle)
  )
df %>%
  lm(modernDance ~ scrapbook, data = .) %>%
  abline()
```
The outlier on the top right makes the line to have a positive trend.


REMOVE OUTLIER 
First, identify the outlier
```{r}
df %>%
  select(state_code, scrapbook) %>%
  filter(scrapbook > 4) %>%  # Select outlier
  print()
```


Bivariate scatterplot without outlier adding fit line
```{r}
df %>%
  select(scrapbook:modernDance) %>% 
  filter(scrapbook < 4) %>%  # filter out outlier
  plot(
    main = "Scatterplot of Google Searches by State",
    xlab = "Searches for \"scrapbook\"",
    ylab = "Searches for \"modern dance\"",
    col  = "gray",  # Color of points
    pch  = 20,      # "Plotting character" (small circle)
  )
df %>%
  filter(scrapbook < 4) %>%  # filter out outlier
  lm(modernDance ~ scrapbook, data = .) %>%
  abline()
```




CLEAN UP 

Clear environment
```{r}
rm(list = ls()) 
```

Clear packages
```{r}
p_unload(all)  # Remove all add-ons
```

Clear plots
```{r}
dev.off()  # But only if there IS a plot
```

Clear console
```{r}
cat("\014")  # ctrl+L
```

Clear mind :)





