---
title: "R Bar plots or bar charts"
output: html_notebook
---

Gráfico de barras: agrupa los valores de una variable respecto a una serie de categorías.

```{r}
data(ChickWeight)
?fivenum # Returns Tukey's five number summary (minimum value, lower-hinge or first quartile, median, upper-hinge or third quartile, maximum value) for the input data.
fiveValues <- fivenum(ChickWeight$weight) # just to see what we're working with
fiveValues
```

```{r}
warnings()
```



```{r}
barplot(fiveValues) # the simplest of bar plots
```

```{r}
barplot(height = fiveValues,
        names.arg = fivenum(ChickWeight$weight),
        horiz = TRUE,
        col = fiveValues,
        main = "Range for Chick Weights"
)
```

```{r}
barplot(height = fiveValues,
        names.arg = fivenum(ChickWeight$weight),
        horiz = FALSE,
        col = fiveValues,
        main = "Range for Chick Weights"
)
```


More on bar charts
pacman must already be installed; then load contributed
packages (including pacman) with pacman
```{r}
pacman::p_load(pacman, tidyverse)
```


Load data to work with:
Using the diamonds dataset from ggplot2
```{r}
?diamonds  # Get information about the diamonds dataset
diamonds   # Display the beginning of the dataset
```



BARPLOT OF FREQUENCIES 
```{r}
?plot     # Get info on "Generic X-Y Plotting
?barplot  # Get info on the "Bar Plots" function
```

Shortest method to make a barplot
```{r}
plot(diamonds$cut)
```


Similar process using pipes
```{r}
diamonds %>%
  select(color) %>%
  plot()
```

Similar code using barplot(); doesn't work
```{r}
diamonds %>%
  select(clarity) %>%
  barplot()  # Error: height must be a vector or a matrix
```

Create frequency vector with table()
```{r}
?table # uses the cross-classifying factors to build a contingency table of the counts at each combination of factor levels.
```

Try again with table()
```{r}
diamonds %>%
  select(clarity) %>%
  table() %>%  # Put data in appropriate format
  barplot()
```

Sort bars by decreasing values (not for ordinal X)
```{r}
diamonds %>%
  select(clarity) %>%
  table() %>%
  sort(decreasing = T) %>%  # Sort table
  barplot()
```


Add options to plot
```{r}
diamonds %>%
  select(clarity) %>%
  table() %>%  # Put data in appropriate format
  barplot(
    main   = "Clarity of Diamonds",
    sub    = "(Source: ggplot2::diamonds)",
    horiz  = T,  # Draw horizontal bars
    ylab   = "Clarity of Diamonds",  # Flip axis labels
    xlab   = "Frequency",
    xlim   = c(0, 15000),  # Limits for X axis
    border = NA,  # No borders on bars
    col    = "#CD0000"  # red3
  )
```



SIDE-BY-SIDE BARPLOT OF FREQUENCIES 

100% stacked bar
```{r}
diamonds %>%
  select(color, clarity) %>%
  plot()
```


Stacked bars: step 1: create table
```{r}
df <- diamonds %>%
  select(color, clarity) %>%
  table() %>%
  print()  # Show table in Console
```


Stacked bars: step 1: create graph
```{r}
df %>%
  barplot(legend = rownames(.))  # Draw plot w/legend
```


Side-by-side bar
```{r}
df %>%
  barplot(
    legend = rownames(.),  # Dot is placeholder for pipe
    beside = T  # Put bars next to each other
  )
```





CLEAN UP 

Clear environment
```{r}
rm(list = ls()) 
```

Clear packages
```{r}
p_unload(all)  # Remove all add-ons
```

Clear plots
```{r}
dev.off()  # But only if there IS a plot
```

Clear console
```{r}
cat("\014")  # ctrl+L
```

Clear mind :)










